---
title: "Tamar_Assignment1"
format: html
editor: visual
---

## Question 1

```{r}
# Load necessary library packages
library(tidyverse)
library(dplyr)

# Import the dataset
avengers <- read_csv("avengers.csv")

# View the dataset to ensure it loaded correctly
view(avengers)

# Check for missing values per variable
missing_values <- colSums(is.na(avengers))

# Display missing values
missing_values

# Check summary statistics for numerical variables
summary(avengers)

# Check for unique values in categorical variables
unique(avengers$superpower)
unique(avengers$north_south)
unique(avengers$died)
```

## Question 2:

Subset the data to include only complete cases:

```{r}
# Remove rows with missing values
avengers_clean <- avengers %>% drop_na()

# Verify no missing values remain
colSums(is.na(avengers_clean))

# Create a new variable: Combat Effectiveness (sum of agility, speed, strength, and willpower)
avengers_clean <- avengers_clean %>%
  mutate(CombatEffectiveness = agility + speed + strength + willpower)

# Check the first few rows to confirm the new variable
head(avengers_clean)

# Summary statistics for Combat Effectiveness
combat_stats <- avengers_clean %>%
  summarise(
    Mean_Combat = mean(CombatEffectiveness, na.rm = TRUE),
    SD_Combat = sd(CombatEffectiveness, na.rm = TRUE),
    Min_Combat = min(CombatEffectiveness, na.rm = TRUE),
    Max_Combat = max(CombatEffectiveness, na.rm = TRUE)
  )

# Display Combat Effectiveness statistics
print(combat_stats)
```

## Question 3:

Subset the data to include only Avengers who did not have a superpower (superpower == "no") and died in battle (died == "yes"):

```{r}
# Subset data for Avengers who died and had no superpowers
avengers_no_superpower_dead <- avengers_clean %>%
  filter(superpower == "no", died == "yes")

# Check the first few rows
head(avengers_no_superpower_dead)

# Save as CSV
write_csv(avengers_no_superpower_dead, "avengers_no_superpower_dead.csv")

# Save as SPSS
library(haven) # package to export to SPSS
write_sav(avengers_no_superpower_dead, "avengers_no_superpower_dead.sav")

```

Summarize key variables (CombatEffectiveness, kills, and injuries) by calculating mean, standard deviation (SD), and range:

Overall summary statistics

```{r}
# Load necessary package
library(dplyr)

# Overall summary for all Avengers without superpowers who died
overall_summary <- avengers_no_superpower_dead %>%
  summarise(
    Mean_Combat = mean(CombatEffectiveness, na.rm = TRUE),
    SD_Combat = sd(CombatEffectiveness, na.rm = TRUE),
    Min_Combat = min(CombatEffectiveness, na.rm = TRUE),
    Max_Combat = max(CombatEffectiveness, na.rm = TRUE),
    
    Mean_Kills = mean(kills, na.rm = TRUE),
    SD_Kills = sd(kills, na.rm = TRUE),
    Min_Kills = min(kills, na.rm = TRUE),
    Max_Kills = max(kills, na.rm = TRUE),
    
    Mean_Injuries = mean(injuries, na.rm = TRUE),
    SD_Injuries = sd(injuries, na.rm = TRUE),
    Min_Injuries = min(injuries, na.rm = TRUE),
    Max_Injuries = max(injuries, na.rm = TRUE)
  )

# Display overall summary statistics
print(overall_summary)
```

Summary statistics by battlefield location

```{r}
# Calculate summary statistics by battlefield location (North vs. South)
battlefield_summary <- avengers_no_superpower_dead %>%
  group_by(north_south) %>%
  summarise(
    Mean_Combat = mean(CombatEffectiveness, na.rm = TRUE),
    SD_Combat = sd(CombatEffectiveness, na.rm = TRUE),
    Min_Combat = min(CombatEffectiveness, na.rm = TRUE),
    Max_Combat = max(CombatEffectiveness, na.rm = TRUE),
    
    Mean_Kills = mean(kills, na.rm = TRUE),
    SD_Kills = sd(kills, na.rm = TRUE),
    Min_Kills = min(kills, na.rm = TRUE),
    Max_Kills = max(kills, na.rm = TRUE),
    
    Mean_Injuries = mean(injuries, na.rm = TRUE),
    SD_Injuries = sd(injuries, na.rm = TRUE),
    Min_Injuries = min(injuries, na.rm = TRUE),
    Max_Injuries = max(injuries, na.rm = TRUE)
  )

# Display battlefield-specific summary statistics
print(battlefield_summary)
```

## Question 4:

-   North had a slightly higher mean Combat Effectiveness (499.78 vs. 491.68). However the South battlefield had a wider spread (higher SD = 189.53 vs. 174.07), meaning individual Avengers had more variability in combat effectiveness.

-   If higher mean Combat Effectiveness is the main indicator, North was slightly more effective. However, the South battlefield had more extreme performers (both weak and strong Avengers).

-   The North battlefield had slightly more injuries (4.60 vs. 4.43), but South had a wider range of injuries.

## Question 5:

-   Kills is the most erroneous variable because it has a very high standard deviation compared to the mean, it has extreme values (e.g., some with 0 kills, others with 79 kills), and it is highly skewed, meaning the mean does not represent most Avengers well.

## Question 6:

Justification 1: Use prior literature/conventions

-   Cohen (1988) suggests that for psychological research, effect sizes of d = 0.2 (small), d = 0.5 (medium), and d = 0.8 (large) are recommended.

-   Prior studies on cognitive differences between superpowered and non-superpowered individuals suggest a medium effect size (d = 0.5).

Justification 2: A priori power analysis

-   A power analysis with Type I error rate set at 0.05, power = 0.80, two-tailed test, and a medium effect size (Cohen's d) = 0.5, can help determine how many participants are needed to detect an effect of interest.

-   This ensures that we have enough participants before conducting the study, avoiding underpowered results.

Now, conduct a priori power analysis to determine sample size:

```{r}
# Load necessary package
library(pwr)

# Define parameters for power analysis
effect_size <- 0.5  # Medium effect size
alpha <- 0.05       # Significance level
power <- 0.80       # 80% power

# Compute required sample size for an independent samples t-test
sample_size <- pwr.t.test(d = effect_size, power = power, sig.level = alpha, type = "two.sample")

# Display required sample size per group
print(sample_size)

```

Check if we have enough participants:

```{r}
# Count number of Avengers in each group
superpower_counts <- table(avengers_clean$superpower)

# Display counts
print(superpower_counts)

```

Conduct independent samples t-test:

```{r}
# Conduct independent samples t-test on IQ by superpower status
t_test_results <- t.test(iq ~ superpower, data = avengers_clean, var.equal = TRUE)

# Display t-test results
print(t_test_results)

```

## Question 7:

I chose a Cohen's d = 0.5 (moderate effect). This suggests that Avengers with superpowers have an IQ score that is about 0.5 standard deviations higher than those without superpowers.

A very large effect (d \> 0.8) would suggest an extreme difference, which may not be realistic. A very small effect (d \< 0.2) would be harder to detect without a large sample size. Choosing d = 0.5 balances realistic detectability and theoretical expectations.

```{r}
# Load necessary package
library(pwr)

# Define chosen effect size (adjust d as needed)
effect_size <- 0.5  # Medium effect

# Compute required sample size for an independent samples t-test
sample_size <- pwr.t.test(d = effect_size, power = 0.80, sig.level = 0.05, type = "two.sample")

# Display required sample size per group
print(sample_size)

```

## Question 8:

```{r}
# Load the necessary package
library(pwr)

# Define power analysis parameters
effect_size <- 0.5   # Medium effect size (Cohen's d)
alpha <- 0.05        # Significance level (Type I error)
power <- 0.80        # Desired power (1 - Type II error)

# Compute the required sample size
sample_size <- pwr.t.test(d = effect_size, 
                          sig.level = alpha, 
                          power = power, 
                          type = "two.sample")

# Display required sample size per group
print(sample_size)

```

## Question 9:

To confirm that there is truly no difference between the groups, an equivalence test or power analysis for detecting a null effect (d = 0) must be conducted.

To test for no meaningful difference, we can define the smallest effect size of interest (SESOI) like d = 0.2. If d \< 0.2, we conclude no meaningful difference exists.

```{r}
# Load TOSTER package
library(TOSTER)

# Compute sample size required for equivalence testing
power_equivalence <- power_t_TOST(n = NULL, 
                                  delta = 0, 
                                  sd = 1, 
                                  eqb = c(-0.2, 0.2),  # Smallest effect size of interest
                                  alpha = 0.05, 
                                  power = 0.80, 
                                  type = "two.sample")

# Display result
power_equivalence

# Count number of Avengers in each group
superpower_counts <- table(avengers_clean$superpower)

# Display counts
print(superpower_counts)
```

## Question 10:

```{r}
# Compute degrees of freedom
n1 <- 780  # Non-superpowered group
n2 <- 32   # Superpowered group
df <- n1 + n2 - 2

# Print degrees of freedom
print(df)

# Load the effectsize package
library(effectsize)

# Compute Cohen’s d from t-value
cohen_d_result <- t_to_d(t = 4.25, df_error = df, paired = FALSE)

# Display Cohen's d and its 95% CI
print(cohen_d_result)

```
